{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-11-30 thesis/landslide.tex
いちおう acknowledge friends

\begin{itemize}
	\item {\tt tell\_landslide\_thread\_switch(int new\_tid)} - to be called during context switch, indicating the newly-running thread
		(must be called with interrupts and/or scheduler preemption disabled)
	\item {\tt tell\_landslide\_sched\_init\_done()} - to be called after scheduler initialization,
		indicating the point after which Landslide should begin analysis.
		Any threads already initialized before this point (init, idle, etc) should be specified with {\tt starting\_threads} (previous subsection).
	\item {\tt tell\_landslide\_forking()} - to be called whenever a new thread is created,
		``immediately'' before the next {\tt thread\_switch()} or {\tt on\_rq()} call for that new thread (i.e., this call sets a flag which the next instance of either of the latter will check to see if the indicated thread is new).
		Most Pebbles kernels will call this twice; once in {\tt fork} and once in {\tt thread\_fork}.
	\item {\tt tell\_landslide\_vanishing()} - to be called whenever a thread ceases to exist,
		``immediately'' before the next {\tt thread\_switch()} or {\tt off\_rq()} call for the exiting thread
		(works similarly to above).
	\item {\tt tell\_landslide\_sleeping()} - to be called whenever a thread is about to {\tt sleep()} waiting for timer interrupts,
		``immediately'' before the next {\tt thread\_switch()} or {\tt off\_rq()} call for the sleeping thread
		(similar to the above).
		Landslide considers sleeping threads to be runnable as normal (they will just take more timer interrupts to arrive at),
		so this call is necessary to distinguish from the case when a thread is descheduled on a non-timer event.
	\item {\tt tell\_landslide\_thread\_on\_rq(int tid)} - to be called when a thread is added to the runqueue
		(must be called with interrupts and/or scheduler preemption disabled).
	\item {\tt tell\_landslide\_thread\_off\_rq(int tid)} - dual of the above.
		If {\tt CURRENT\_THREAD\_\allowbreak{}LIVES\_ON\_RQ=0} (described above), this should be invoked (among other times) during context switch with the TID of the thread about to start running.
		Alternatively (thanks sully), even for a kernel which takes the current thread off its literal runqueue,
		the annotator may use these two calls to indicate the ``abstract runqueue'' which includes the current thread as well,
		and set {\tt CURRENT\_THREAD\_LIVES\_ON\_RQ=1}.
\end{itemize}

>>>>

\begin{itemize}
	\item {\tt tell\_landslide\_thread\_switch(int new\_tid)} - to be called during context switch, indicating the newly-running thread
		(must be called with interrupts and/or scheduler preemption disabled)
	\item {\tt tell\_landslide\_sched\_init\_done()} - to be called after scheduler initialization,
		indicating the point after which Landslide should begin analysis.
		Any threads already initialized before this point (init, idle, etc) should be specified with {\tt starting\_threads} (previous subsection).
	\item {\tt tell\_landslide\_forking()} - to be called whenever a new thread is created,
		``immediately'' before the next {\tt thread\_switch()} or {\tt on\_rq()} call for that new thread (i.e., this call sets a flag which the next instance of either of the latter will check to see if the indicated thread is new).
		Most Pebbles kernels will call this twice; once in {\tt fork} and once in {\tt thread\_fork}.
	\item {\tt tell\_landslide\_vanishing()} - to be called whenever a thread ceases to exist,
		``immediately'' before the next {\tt thread\_switch()} or {\tt off\_rq()} call for the exiting thread
		(works similarly to above).
	\item {\tt tell\_landslide\_sleeping()} - to be called whenever a thread is about to {\tt sleep()} waiting for timer interrupts,
		``immediately'' before the next {\tt thread\_switch()} or {\tt off\_rq()} call for the sleeping thread
		(similar to the above).
		Landslide considers sleeping threads to be runnable as normal (they will just take more timer interrupts to arrive at),
		so this call is necessary to distinguish from the case when a thread is descheduled on a non-timer event.
	\item {\tt tell\_landslide\_thread\_on\_rq(int tid)} - to be called when a thread is added to the runqueue
		(must be called with interrupts and/or scheduler preemption disabled).
	\item {\tt tell\_landslide\_thread\_off\_rq(int tid)} - dual of the above.
		If {\tt CURRENT\_THREAD\_\allowbreak{}LIVES\_ON\_RQ=0} (described above), this should be invoked (among other times) during context switch with the TID of the thread about to start running.
		Alternatively, %(thanks sully),
		even for a kernel which takes the current thread off its literal runqueue,
		the annotator may use these two calls to indicate the ``abstract runqueue'' which includes the current thread as well,
		and set {\tt CURRENT\_THREAD\_LIVES\_ON\_RQ=1}.
\end{itemize}

### 2018-11-30 thesis/thesis.tex
いちおう acknowledge friends

\begin{acknowledgments}
I honor here the many people who supported me during these 7 years and made this journey possible.

@@

Professor Garth Gibson is largely responsible for shaping me into the researcher I am today.
%Thanks to Garth's guidance,
%this thesis has grown into something that past-me,
%first starting research with the simple vision ``make a better autograder for 410'',
%could not have then imagined.
Despite having primary research interests in other fields,
Garth enthusiastically took on the project,
and always pushed me to explore new problem domains,
to design scientifically thorough experiments,
to explain difficult concepts approachably,
to appreciate related work as charitably as possible,
and to seek guidance from industry people with relevant experience,
despite my often stubborn refusal to leave my little comfort zone.
So, even when it took me sending 10+ email reminders to get your attention,
truly,
thanks for everything, Garth.

Professor David A. Eckhardt
provided unending guidance on education, writing, career, life, you name it,
helped revise each semester's iteration of the recruiting lecture with a keen eye for the student mind,
contributed substantially to the design of the survey,
and supplied constant enthusiasm and encouragement that Landslide was a boon to 15-410.
I can't express my gratitude to Dave strongly enough,
even if it took until just this week for half of the Landslide-friendly tests to be included in the official P2 test suite.
%even though three of the Landslide-friendly tests never made it upstream into the official P2 test suite.

Ji\v{r}\'{i} \v{S}im\v{s}a, another student of Garth's,
first sparked my interest in the field,
and then gave invaluable support during my first few years.
Ji\v{r}\'{i}'s mentorship
oriented me in the world of research and helped me find my bearings as a grad student,
from teaching me how DPOR works
to helping revise my paper drafts
to offering the kind of candid career advice you can't get from a professor.
Had I not met Ji\v{r}\'{i} and learned about dBug during grad school applications way back,
none of this would have happened.

{Many others have collaborated with me directly on the research itself,
whom I thank here in approximate chapter order.
Michael J. Sullivan and Nathaniel Filardo were Landslide's first users besides myself back in its prototype days.
Joshua Wise, Miriam Zimmerman, and Glenn Willen helped revise my explanation of DPOR in Chapter~\ref{chap:background}.
Professor Brandon Lucia provided valuable comments and revisions to Chapter~\ref{chap:quicksand}'s content
that ultimately led to its publication at OOPSLA 2016.
Michael J. Sullivan double-checked the proofs therein.
Professor Timothy Zhu from Penn State University
and Professor Haryadi Gunawi and TA Kevin Zhao from the University of Chicago
graciously allowed me to use their students as guinea pigs
and helped greatly to improve Landslide's stability and robustness for use beyond CMU's walls,
greatly enriching Chapter~\ref{chap:education}'s evaluation.
David Blum and Skye Toor provided
invaluable advice on statistical analysis
that helped me evaluate the student
% justify last line bc of manual page break
\unskip\parfillskip 0pt \par}

@@

The community of 15-410 TAs,
having surrounded me pretty much since starting undergrad,
nurtured my love of teaching
well before taking the class myself and well after graduation.
As skilled as the instructor is,
15-410 could not be the world's best operating systems class
without TAs who value thorough grading, the Socratic method,
and empathy to struggling students as much
from year to year as the crew I have known.
It was a blessing to serve, and later do research, among you all.
Thanks also to Wind River for generously providing CMU with Simics educational licenses.
I am also grateful to Adam Blank and Miriam Zimmerman
for founding and organizing the class Great Practical Ideas,
which teaches oft-neglected programming skills such as version control and debugging,
which I had the opportunity to teach as head instructor one semester.

I thank the Parallel Data Lab, my research group at CMU,
for supporting me at the PDL retreat every fall,
pushing me to refine my presentation skills and to network with strangers (now familiar faces) from industry.
I also thank those industry attendees from each of the PDL Consortium member companies
for their interest, even in more theoretical work,
and the enlightening conversations resulting therefrom.
I thank everyone outside my immediate research circle who,
whether at the retreat or otherwise,
insisted how cool my research was for directly helping students --
every such reminder helped me keep my head up during darker times.

I am grateful to the ThursDz Council,
my group of now-former grad student, Shadyside-bound (in spirit if not in person) restaurant-hoppers,
for both friendship and mentorship over the years.
In particular,
%in alphabetical for lack of a better order,
Chris Martens,
Jason Reed,
Jim McCann,
Rob Simmons,
Tom Murphy 7,
and
William Lovas:
you each have been role models to me
perhaps more than you know.

@@

My parents, Eve and David Blum, and my grandparents, Margie Granach and Elsie Blum,
provided bottomless emotional support throughout my education.
I am grateful beyond words for the privileged life they provided
that allowed me the opportunity to pursue higher education.

% [ ] csd bgames
% [ ] undergrad frenz
% [ ] new frenz
% [ ] family, incl louie

Lastly,
the Android: Netrunner community has been such an important part of my life
for nearly as long as I've been a grad student.
Netrunner is a tabletop strategy deckbuilding card game
that depicts cyberspace hacking battles (completely different from real-life computer security)
alongside dystopian social issues (some of which seem oddly more familiar with each passing year).
Its characters have a diversity of ethnicity, gender identity, and sexual orientation
unparalleled in any other game I know of,
and the community that has grown around it reflects those values.
I thank in particular the other Stimhack moderators,
the members of {\sf \#rainbow-coalition},
and the local Pittsburgh crew
for their support and encouragement
and friendships very real and deep, even when forged only in text,
%\footnote{and Slack emojis},
that will long outlive the silly card game that brought us together.
%It'd be an injustice to list only a few names here and omit others, but you know who you are.

>>>>

\begin{acknowledgments}
\revision{I honor here the many people who supported me during these 7 years and made this journey possible.

@@

Professor Garth Gibson is largely responsible for shaping me into the researcher I am today.
%Thanks to Garth's guidance,
%this thesis has grown into something that past-me,
%first starting research with the simple vision ``make a better autograder for 410'',
%could not have then imagined.
Despite having primary research interests in other fields,
Garth enthusiastically took on the project,
and always pushed me to explore new problem domains,
to design scientifically thorough experiments,
to explain difficult concepts approachably,
to appreciate related work as charitably as possible,
and to seek guidance from industry people with relevant experience,
despite my often stubborn refusal to leave my own little comfort zone.
So, even when it took me sending 10+ email reminders to get your attention,
truly,
thanks for everything, Garth.

Professor David A. Eckhardt
provided unending guidance on education, writing, career, and life in general, %life, you name it,
helped revise each semester's %iteration of the
recruiting lecture with a keen eye for the student mind,
contributed several clever survey questions, %substantially to the design of the survey,
and supplied constant %enthusiasm and
encouragement about Landslide's value to 15-410, %that Landslide was a boon to 15-410,
not to mention the red-ink-encrusted printed draft of this document I received immediately after my defense.
My debt of thanks to Dave is of such magnitude that I may never hope to pay it back, but perhaps forward instead.
%I can't express my gratitude to Dave strongly enough,
%even if it took until just this week for half of the Landslide-friendly tests to be included in the official P2 test suite.
%even though three of the Landslide-friendly tests never made it upstream into the official P2 test suite.

Ji\v{r}\'{i} \v{S}im\v{s}a, another former student of Garth's,
first sparked my interest in the field,
and then gave invaluable support during my first few years.
Ji\v{r}\'{i}'s mentorship
oriented me in the world of research and helped me find my bearings as a grad student,
from teaching me how DPOR works
to helping revise my paper drafts
to offering the kind of candid career advice you can't get from a professor.
Had I not met Ji\v{r}\'{i} and learned about dBug during grad school applications way back,
none of this would have happened.

{Many others have collaborated with me directly on the research itself,
whom I thank here in approximate chapter order.
Michael J. Sullivan and Nathaniel Wesley Filardo were Landslide's first users besides myself back in its prototype days.
Joshua Wise, Miriam Zimmerman, and Glenn Willen helped revise my explanation of DPOR in Chapter~\ref{chap:landslide}.
Professor Brandon Lucia provided valuable comments and revisions to Chapter~\ref{chap:quicksand}'s content
that ultimately led to its publication at OOPSLA 2016.
Michael J. Sullivan double-checked the proofs therein.
Professor Timothy Zhu from Penn State University
and Professor Haryadi Gunawi and TA Kevin Zhao from the University of Chicago
graciously allowed me to use their students as guinea pigs
and helped greatly to improve Landslide's stability and robustness for use beyond CMU's walls,
greatly enriching Chapter~\ref{chap:education}'s evaluation.
David Blum and Skye Toor provided
invaluable advice on statistical analysis
that helped me evaluate the student
% justify last line bc of manual page break
\unskip\parfillskip 0pt \par}

@@

The community of 15-410 TAs
%having surrounded me pretty much since starting undergrad,
nurtured my love of teaching
well before taking the class myself and well after graduation.
%As skilled as the instructor is,
15-410 could not be the world's best operating systems class
without %TAs
course staff %(staves)
who value thorough grading, the Socratic method,
and empathy to struggling students as much
from year to year as the crew I have known.
It was a blessing to serve, and later to do research, among you all.
Thanks to Wind River for generously
providing CMU with Simics educational licenses.
%I am also grateful
Thanks also to Adam Blank and Miriam Zimmerman
for founding and organizing the class Great Practical Ideas,
which teaches oft-neglected programming skills such as version control and debugging,
which I had the opportunity to teach as head instructor one semester.

I thank the Parallel Data Lab, my research group at CMU,
for supporting me at the PDL retreat every fall,
pushing me to refine my presentation skills and to network with strangers (now familiar faces) from industry.
I also thank those %industry
attendees from each of the PDL Consortium member companies
for their interest %, even in more theoretical work,
and for the enlightening conversations resulting therefrom.
%Thanks to the anonymous OOPSLA reviewer who went the extra mile to read my supplementary proofs
%and urged me to improve the formalism to be more foundational
%(the lack of any stronger criticism a silent vote of confidence for their overall correctness).
Thanks to Niko Matsakis for the mentorship and collaboration during my two summers as a Rust research intern at Mozilla,
and to the rest of the Rust community as well for their extremely important work
bringing better concurrent programs to the world from a different angle.
I thank everyone outside my %immediate
research circle who
%whether at the retreat or otherwise,
insisted how cool my research was for directly helping students --
every such reminder helped me keep my head up during darker times.

%I am grateful to
The ThursDz Council,
my group of now-former grad student, Shadyside-bound (in spirit if not in person) restaurant-hoppers,
surrounded me with both friendship and mentorship over the years.
In particular,
%in alphabetical for lack of a better order,
Chris Martens,
Jason Reed,
Jim McCann,
Rob Simmons,
Tom Murphy 7,
and
William Lovas:
you each have been role models to me
perhaps more than you know.

@@

My parents, Eve and David Blum, and my grandparents, Margie Granach and Elsie Blum,
provided bottomless emotional support throughout my education.
I appreciate beyond words for the privileged life they provided
that allowed me the opportunity to pursue higher education.

Looking back on my 11 years at CMU,
I am especially grateful for the friendships made early on %in my undergrad years
that continued well past first graduation.
From those who stayed in Pittsburgh %with me
to those who %moved away but
kept in touch,
from my regular tabletop, board, and/or dance gaming partners
to those who %inexhaustibly
welcomed me with %company and
a couch to sleep on during travel,
I thank
%-- and in any order other than alphabetical here would be far too cruel --
Adam Blank,
Alan Vangpat,
Alex Yuschik,
Anand Subramanian,
Andrew Krieger,
Car Bauer,
Carolyn Sawyer,
Elizabeth Kemp,
Elly Fong-Jones,
Emma Cating-Subramanian,
Eric Faust,
Gabe Routh,
Greg Hanneman,
Hazel Vird\'{o},
Jack Ferris,
Jake Lengyel,
Jason Deng,
Josh Keller,
Joshua Wise,
Josiah Boning,
Julia Tuttle,
Kellie Medlin,
Laura Abbott,
Margaret Meyerhofer,
Michael ``rntz'' Arntzenius,
Michael J. ``Sully'' Sullivan,
Miriam Zimmerman,
Naomi Saphra,
Nathaniel Wesley Filardo,
Philip ``8'' Gianfortoni,
Rhett Lauffenburger,
Richard Ha,
Rory Glenn,
Ryan Pearl,
and
Todd Eisenberger
for their wonderful friendship.
I am equally grateful to the new friends I made during my graduate years,
whether through board/card/dance gaming, tea drinking, %and/or
movie watching,
and/or as more permanent Pittsburgh residents that I am extremely torn up about moving away from:
%leaving behind (for now at least):
Alexandra Lee Falk,
Alexis Dyer,
Barbara Jensen,
Brendan McShane,
Brian Saghy,
Carlo Angiuli,
Cassie Orr,
Dan Guzek,
Danny Gratzer,
David Renshaw,
Emily Forney,
Evan Cavallo,
Gabriele Farina,
Grant Wu,
Jenny Lin,
Karl Schulz,
Kristy Gardner,
Lea Albaugh,
Maija Mednieks,
Marlena Abraham,
Matt Stanec,
Matthew Maurer,
Sarah R. Allen,
Sol Boucher,
Stefan Muller,
Zachary Waldman,
Ziv Scully,
and the rest of ThursDz already named previously.
And to anyone who may read this list
and feel even slightly disappointed that their name was left off:
I owe you a nice dinner. Get in touch.

日本語を
一緒に
勉強したり、
会話したり、
色々教えてくださったりした
学習者さんたちとネイティブスピーカーさんたちへ、
心から感謝を申し上げます。
外国語を学ぶのは一生の二番目の達成感かもしれません。

Lastly,
the Android: Netrunner community has been such an important part of my life
for nearly as long as I've been a grad student.
Netrunner is a deckbuilding strategy card game
that depicts cyberspace hacking battles (completely different from real-life computer security)
alongside dystopian social issues (becoming strangely more familiar with each passing year).
Its characters have a diversity of ethnicity and gender identity %and sexual orientation
unparalleled in any other game I know of,
and the community that has grown around it reflects those values.
I thank especially the other Stimhack moderators,
the members of {\sf \#rainbow-coalition},
and the local Pittsburgh crew
for their support and encouragement and
friendships very real and deep, even when forged only in text,
%\footnote{and Slack emojis},
that will long outlive the silly card game that brought us together.
%It'd be an injustice to list only a few names here and omit others, but you know who you are.
}
